---
title: "Exploratory Analysis"
author: "Aaron Palumbo"
date: "Sunday, July 05, 2015"
output: html_document
---

## Background Information

From the data dictionary:
>The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain. See 'features_info.txt' for more details. 

This means that we are not dealing with time domain vectors. Each observation represents 561 measurements of a 2.56 second window. We will attempt to ascertain what activity the user was engaged in from this 2.56 second window.


## Environment / Dependencies

### Dependencies
```{r dependencies, warning=FALSE, message=FALSE}
library(knitr)
# Suppress messages and warnings in all chuncks
opts_chunk$set(message=FALSE, warning=FALSE)

# Libraries
library(caret)
library(randomForest)

# Processing
library(doParallel)
registerDoParallel(cores = 4)

```


### Working Directory / File Paths

```{r environment}
## Working Directory
proj_dir <- "IS624_Final_Project"
if(basename(getwd()) == proj_dir){
  setwd("./code")
}

if(!(basename(getwd()) == "code")){
  break
}

## File Paths
dataDir <- "../data"
trainDataDir <- file.path(dataDir, "UCI HAR Dataset", "train")
testDataDir <- file.path(dataDir, "UCI HAR Dataset", "test")
```

### Custom Functions

```{r custom_functions}

```


## Data Partitioning

According to the file README.txt, the data was randomly partitioned into two sets where 70% of the volunteers were selected for generating the training data and 30% the test data. Splitting the data this way, by volunteer, is  a good idea in that we want our algorithm to be general enough to work from person to person, not just be good at identifying activities of a particular person. This would offer a good starting point for a device of this type. After it is purchased, it might make sense to tune the parameters to the specific person.


## Load Training Data

```{r train_data}
## Sensor Data
df.sns <- read.csv(file.path(trainDataDir, "X_train.txt"), 
                     header=FALSE, sep="")
snsColNums <- 1:ncol(df.sns)

features <- readLines(file.path(dataDir, "UCI HAR Dataset", "features.txt"))

## Subject ID
sub <- as.factor(readLines(file.path(trainDataDir, "subject_train.txt")))

## Activity ID
act <- as.factor(readLines(file.path(trainDataDir, "y_train.txt")))
```


```{r test_data}
## Sensor Data
df.sns.test <- read.csv(file.path(testDataDir, "X_test.txt"), 
                        header=FALSE, sep="")
# snsColNums <- 1:ncol(df.sns)

## Subject ID
sub.test <- as.factor(readLines(file.path(testDataDir, "subject_test.txt")))

## Activity ID
act.test <- as.factor(readLines(file.path(testDataDir, "y_test.txt")))
```


## EDA

### Sanity Checks

```{r sanity_checks}
## All sensor columns should contain numeric data
# TRUE
all(sapply(df.sns[1, ], class) == "numeric")

## Activity vectory should be a factor
class(act)

## Subject should be a factor
class(sub)

## Counts of activities
# Each activity is well represented
table(act)

## Counts of subjects
# Each subject is  well represented
table(sub)
```

### Variable Information

```{r}
## Correlated Variables
cor.mat <- cor(df.sns)
corFilter <- findCorrelation(cor.mat)

df.sns.removeCor <- df.sns[ ,-corFilter]
cor.mat2 <- cor(df.sns.removeCor)
# corrplot(cor.mat2)
```

### Random Forest


#### Feature Selection

```{r}
# use RFE from caret for feature selection
subsets <- c(2, 9, 35, seq(120, 300, 30), 400, 561)
# subsets <- c(2, 9)

rfRFE <- list(summary=defaultSummary,
              fit=function(x, y, first, last, ...){
                library(randomForest)
                randomForest(x, y, importance = first, ...)
                },
              pred = function(object, x) predict(object, x),
              rank = function(object, x, y) {
                vimp <- varImp(object)
                vimp <- vimp[order(vimp$Overall, decreasing=TRUE),, drop=FALSE]
                vimp$var <- rownames(vimp)
                vimp
                },
              selectSize = pickSizeBest,
              selectVar = pickVars)

ctrl <- rfeControl(functions=rfFuncs,
                   returnResamp="all",
                   method="repeatedcv",
                   number=10,
                   repeats=5,
                   verbose=FALSE,
                   allowParallel=TRUE)

# s <- sample(1:7352, 1000)
rfProfile <- rfe(df.sns, act, sizes=subsets, rfeControl=ctrl)
save(rfProfile, file='rfProfile')
```

```{r}
subsets <- seq(14, 35, 7)

trControl <- trainControl(method="cv", number=5)

nnetGrid <- expand.grid(.decay = c(0, 0.01, 0.1),
                        .size = c(5:10),
                        .bag = FALSE)

ctrl.nn <- rfeControl(funcitons=caretFuncs,
                   returnResamp="all",
                   method="cv",
                   number=5,
                   verbose=FALSE,
                   allowParallel=TRUE)

coltouse <- rownames(vimp)[1:35]

rfP.nn <- rfe(df.sns[,coltouse],  act, sizes=subsets, rfeControl=ctrl.nn,
              trControl=trControl, tuneGrid=nnetGrid, method="avNNet",
              preProc = c("center", "scale"), linout = TRUE,
              trace=FALSE, MaxNWts = 371, maxit = 500)
```


All predictors

```{r randomForest}
if("fit.rf" %in% list.files()){
  load("fit.rf")
}else{
  fit.rf <- randomForest(x=df.rf, y=act)
  save(fit.rf, file="fit.rf")
}

```

Remove correlated
```{r}
fit2.rf <- randomForest(act ~ ., data=cbind(df.sns.removeCor, act))
fit2.rf <- randomForest(x=df.sns[ ,-corFilter], y=act)
```

```{r}
fit.import <- varImp(fit.rf)
plot(fit.import$Overall)
fit.import.order <- order(fit.import$Overall, decreasing = TRUE)

fit3.rf <- randomForest(act ~ ., data=cbind(df.imp200, act))
```

```{r}
fitControl <- trainControl(method="repeatedcv",
                           number=10,
                           repeats=10)
train.rf <- train(x=df.sns[ ,  fit.import.order[1:100]], y=act,
                  method="rf",
                  trControl = fitControl)

```

Initial experimenting shows that better performance can be achieved by reducing the number of predictors. So far the top 200 as selected by top varImp in full model is best.

There seem to be some highly correlated variables as well that might reduce performance.



```{r rfcv}
if("rfcv.result" %in% list.files()){
  load("rfcv.result")
}else{
  rfcv.result <- rfcv(dn.sns, act)
  rfcv.result <- rfcv(df.sns, act, cv.fold=5, scale="log", step=0.1, recursive=TRUE)
  save(rfcv.result, file="rfcv.result")
}

```

```{r}
result <- replicate(5, rfcv(df.sns, act, scale="log", step=0.4, recursive=TRUE), simplify=FALSE)
error.cv <- sapply(result, "[[", "error.cv")
error.cv
```



```{r}
n <- 100
p <- 40
sigma <- 1
set.seed(1)
sim <- mlbench::mlbench.friedman1(n, sd = sigma)
colnames(sim$x) <- c(paste("real", 1:5, sep = ""),
                     paste("bogus", 1:5, sep = ""))
bogus <- matrix(rnorm(n * p), nrow = n)
colnames(bogus) <- paste("bogus", 5+(1:ncol(bogus)), sep = "")
x <- cbind(sim$x, bogus)
y <- sim$y
```

